Texture Mapping Indoor Environments with Imperfect Camera Localization

### 2012-06-20 3dimpvt/paper.tex
sending draft to prof

%%%%%%%%% TITLE
\title{\LaTeX\ Author Guidelines for 3DIMPVT Proceedings}

@@

The first step in automated 3d modeling is the physical scanning of
the environment's geometry. An indoor modeling system must be able to
calculate camera locations within an environment while simulatenously
reconstructing the 3D structure of the environment itself. This
problem is studied by the robotics and computer vision communities as
the simultaneous localization and mapping (SLAM) problem, and is
generally solved using a combination of laser range scanners, cameras,
and inertial measurement units (IMUs).

@@

\section{Backpack Modeling System}
Human-operated data acquisition systems provide unique advantages over
vehicular-mounted systems in terms of agility and
portability. Unfortunately, human-operated systems also suffer from a
lack of automation and stability, resulting in much higher data
variance and localization error. As a result, common methods for
texture mapping generally produce poor results, as shown in section
3. Before discussing how to overcome these challenges, we will first
provide an overview of the backpack modeling system from which our
test data was obtained.

\subsection{Data Acquisition Hardware}
The backpack modeling system that captured our data contains five 2D
laser range scanners, two cameras, an orientation sensor, and an
IMU. The laser scanners are mounted orthogonally and have a 30-meter
range and a 270$^{\circ}$ field of view. The two cameras are equipped
with fisheye lenses, reaching an approximately 180$^{\circ}$ field of
view, and are mounted with one facing left and one facing right. These
cameras take images at the rate of 5 Hz. The orientation sensor
provides orientation parameters at a rate of 180 Hz. The IMU provides
highly accurate measurements of all 6 DOF at 200 Hz, and is used as a
ground truth reference.

@@

\begin{figure}
  \centering
  \includegraphics[height=2in]{Projection.pdf}
  \caption{Planes are specified in 3D space by four corners $C_1$ to
    $C_4$. Images are related to each plane through the camera
    matrices $P_{1..M}$. }
  % \caption{The plane is specified in 3D space by the four corners
  % $C_1$ to $C_4$. Images are related to the plane through the camera
  % matrices $P_{1..M}$. For example, the 3D point X in the world
  % coordinate system is related to the image point x shown in the
  % figure as $x = P_i X$.}
  \label{fig:projection}
\end{figure}

In all subsequent sections, we will discuss the process of texture
mapping a single plane, as the texturing of each of our planes is
completely independent and can be completed in parallel.

The geometry of the texture mapping process for a plane is shown in
Figure 1.  As described in the previous section, we are provided with
a set of $M$ images with which we must texture our target plane. Each
image has a camera matrix $P_i$ for $i=1..M$, which translates a 3D
point in the world coordinate system to a 2D point or pixel in image
$i$'s coordinates. If the 3D world point is not contained in the
image, the 2D point will simply be outside of the image boundaries. A
camera matrix $P_i$ is composed of the camera's intrinsic parameters,
such as focal length and image center, as well as extrinsic parameters
which specify the rotation and translation of the camera's position in
3D world coordinates at the time that image $i$ was taken. These
extrinsic parameters are determined by the backpack hardware and
localization algorithms mentioned in Section 2. A point $X$ on the
plane in 3D space can be related to its corresponding pixel $x$ in
image $i$ through the following equation:

@@

Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
we can texture the plane by discretizing it into small square tiles,
generally about 5 pixels across, and picking an image to texture each
tile with. We choose to work with rectangular units to ensure that
borders between any two distinct images in our final texture will be
either horizontal or vertical. Since most environmental features
inside buildings are horizontal or vertical, any seams in our texture
will intersect them miniimally and be less noticeable.

In order to select an image for texturing tile $t$, we must first
gather a list of candidate images that contain all four of its
corners, which we can quickly check using the projection method
above. Furthermore, each candidate image must have been taken at a
time when its camera had a clear line-of-sight to the center of $t$,
which can be calculated using standard ray-polygon intersection tests
between the camera location, our section, and other planes, all in
world coordinates.

@@

Similar to a caching mechanism, we select the best image for a tile
$t$ by searching through two subsets of images for a good candidate
before searching through the entire set. The first subset of images is
the set of images selected by adjacent tiles that have already been
processed. We must first check which images can map to $t$, and then
of those, we make a choice according to the same scoring function in
figure \ref{fig:scoringFunction}. Rather than blindly reusing this
image, we ensure it meets a predefined score threshold to be
considered a good image. If no good image is found, we then check our
second subset of images, which consists of images that were taken near
the images in the first subset, both spatially and temporally. These
images are not the same as the ones used for neighboring tiles, but
they were taken at a similar location and time, suggesting that their
localization and projection will be very similar. Again, if no good
image is found according to a threshold, we then must search the
entire set of candidate images.

@@

The results of this caching approach are shown above in figure
\ref{fig:caching}.  Discontinuities have been reduced overall, but the
amount of remaining seams suggests that image selection alone cannot
produce seamless textures. The image projections themselves will have
to be adjusted in order to reliably produce clean textures.

@@

Rather than perform camera or image adjustments across many thousands
of images, we choose to work with the more limited set of images
corresponding to those chosen by the direct mapping approach, without
caching. This set of images constitutes a good candidate set for
generating a final seamless texture, since it meets three important
criteria. First, this set of images contains at least one image that
covers each tile on our plane. Thus, unless extreme changes in
localization occur, we can ensure that there will be no holes in our
final texture. Second, since images are all selected according to the
same scoring function in figure \ref{fig:scoringFunction}, we know
that our images are all taken at a head-on angle and should project
onto the plane in similar ways. Third, as a side result of the scoring
function, selected images are only good candidates for the tiles near
their center of projection. Thus, there should be plenty of overlap
between selected images, allowing for some degree of shifting as well
as blending between them. With this set of images, we now examine two
approaches towards refining and combining their projections, before
demonstrating our own.

@@

It should be evident that our localization adjustments resulted in a
major improvement. A great strength of our tiling approach for image
selection is that it extracts the areas with good camera angles from
each image, and thus can handle images of all types. As demonstrated
on the floor plane in figure \ref{fig:floor_suboptimal}, we can run
this process even with extremely poor image projections, where the
amount of quality texture for each image is limited. Additionally,
since our plane is discretized, we can easily handle arbitrarily
complex planes as well.

@@

\subsection{Texture Mapping with Seam Minimization}
The cached tile approach selected images based on individual and
neighboring quality of image selections, in an effort to reduce seams
between tiles. In a more optimal case, as the backpack system provides
for walls, images are all taken from close distances and have
near-rectangular projections and thus a great score according to the
scoring function in figure \ref{fig:scoringFunction}. Thus, we can
reason that rather than selecting the set of best images, since all
images are good, we should instead select the best set of images, such
that the selection together results in the cleanest final texture. We
will accomplish this by using entire images where possible, defining a
cost function between images, and seeking to minimize our total cost.

@@

Fortunately, by virtue of our indoor environments, the vast majority
of surface geometry is either horizontal or vertical, with high
amounts of right angles. This means that after masking out occluded
areas, our image projections will remain largely rectangular. We can
thus recursively split each image into rectangular pieces, and perform
occlusion checks, similar to the process done for tiles earlier. To
actually occlude out rectangles, we simply remove their texture, as we
will ensure that untextured areas are never chosen for texture
mapping.

\subsubsection{Inter-Image Cost Functions}
In order to objectively decide which set of images results in the
cleanest texture, we need a cost function such that we can evaluate
the visibiilty of seams between images in our set. A straightforward
cost function that accomplishes this is the sum of squared pixel
differences in overlapping regions between all pairs of
images. Minimizing this cost function encourages image boundaries to
occur in featureless areas, such as bare walls, as well as in areas
where images match extremely accurately.

Another possible cost function is overall edge energy, i.e. the sum of
the smoothed gradient over seams. Minimizing this encourages image
boundaries to be placed in featureless areas even more than the first
cost function. For the results shown throughout this paper, the first
cost function was used, as the regular occurrence of featureless areas
was not guaranteed in our datasets, and our image matches do in fact
match at the pixel level in many cases.

@@

Now that we have a cost function defined, we mechanically select the
set of images for which the overall cost function is minimized. Since
we aim to cover the entirety of our plane, our problem is to minimally
cover a polygon(our plane), using other polygons of arbitrary geometry
(our image projections), with the added constraint of minimizing our
cost function between chosen images. This is a complex problem, though
we can take a number of steps to simplify it. Returning again to the
optimality of our situation when texturing walls however, we can make
a quick simplification for the sake of efficiency. Given that our
wall-texture candidate images were all taken from a head-on angle, and
assuming only minor rotations were made during localization
refinement, we can reason that their projections onto the plane are
approximately rectangular. By discarding the minor excess texture and
cropping them all to be rectangular, our problem becomes the
conceptually simpler problem of filling a polygon with rectangles,
such that the sum of all edge costs between each pair of rectangles is
minimal. We thus also retain axis-aligned image boundaries, along with
their advantages explained in section 3.

By taking one more shortcut we can simplify our problem down even
further. Since this approach is already assuming optimal images, we
know it will be applied only on walls, which are the focus of the
backpack modeling system. For wall planes, care is taken so that
images contain the entirety of their floor-to-ceiling range, and thus
we do not have wall images such that one should be projected
vertically above the other. In essence, we need only to ensure
horizontal coverage of our planes, as our images provide full vertical
coverage themselves. We can thus construct a Directed Acyclic Graph
(DAG) from the images, again with edge costs defined by our cost
function above, and solve a simple shortest path problem to find an
optimal subset of images with regard to the cost functions.

@@

In rare cases where the vertical dimension of the plane is not
entirely covered by a chosen image, we are left with a hole where no
image was chosen to texture. Rather than reverting to a 2D-coverage
problem, we can elect to simply fill the hole by selecting images in a
greedy fashion with respect to the same cost function.

@@

Although our preprocessing steps and image selections in either method
attempt to minimize all mismatches between images, there are cases,
where due to different lighting conditions or inaccuracies in geometry
or projection, where we simply have unavoidable discontinuities. These
can however be treated and smoothed over by applying alpha blending
over image seams.  Whether the units we are blending are
rectangularly-cropped images or rectangular tiles, we can apply the
same blending procedure, as long as we have a guaranteed overlap
between units.

@@

\section{Conclusion and Results}
In this paper, we have shown how to accurately texture map models,
even when provided with imprecise camera localization data. We have
generalized one approach to any manner of planes and images, and
successfully textured both simple rectangular walls as well as complex
floor and ceiling geometry. We have also presented an optimized
texturing method that takes advantage of our localization refinement
process and produces cleaner textures on planes where multiple head-on
images are available. Ceilings and floors textured with the robust
approach, and walls textured with the optimized approach, are
displayed in figure \ref{fig:results}. A more detailed walkthrough
demonstrating fully textured 3D models using the approaches in this
paper is available in the accompanying video to this paper.

>>>>

%%%%%%%%% TITLE
\title{Texture Mapping Indoor Environments with Imperfect Camera Localization}

@@

@@

\section{Backpack Modeling System}
The first step in automated 3D modeling is the physical scanning of
the environment's geometry. An indoor modeling system must be able to
calculate camera locations within an environment while simulatenously
reconstructing the 3D structure of the environment itself. This is
known as the simultaneous localization and mapping (SLAM) problem, and
is generally solved by taking readings from laser range scanners,
cameras, and inertial measurement units (IMUs) at multiple locations
within the environment.

Mounting such devices on a human-carried platform provides unique
advantages over vehicular-based systems in terms of agility and
portability. Unfortunately, human-operated systems also suffer from a
lack of automation and stability, resulting in much higher data
variance and localization error. As a result, common methods for
texture mapping generally produce poor results, as later shown in
section 3. Before discussing how to overcome these challenges, we will
first provide an overview of the backpack modeling system from which
our test data was obtained.

\subsection{Data Acquisition Hardware}
The backpack-mounted modeling system that captured our data contains
five 2D laser range scanners, two cameras, an orientation sensor, and
an IMU. The laser scanners are mounted orthogonally and have a
30-meter range and a 270$^{\circ}$ field of view. The two cameras are
equipped with fisheye lenses, reaching an approximately 180$^{\circ}$
field of view, and are mounted with one facing left and one facing
right. These cameras take images at the rate of 5 Hz. The orientation
sensor provides orientation parameters at a rate of 180 Hz. The IMU
provides highly accurate measurements of all 6 DOF at 200 Hz, and is
used as a ground truth reference for localization processes.

@@

\begin{figure}
  \centering
  \includegraphics[height=2in]{Projection.pdf}
  \caption{Planes are specified in 3D space by four corners $C_1$ to
    $C_4$. Images are related to each plane through the camera
    matrices $P_{1..M}$. }
  \label{fig:projection}
\end{figure}

In all subsequent sections, we will discuss the process of texture
mapping a single plane, as the texturing of each of our planes is
independent and can be completed in parallel.

The geometry of the texture mapping process for a plane is shown in
figure \ref{fig:projection}.  As described in the previous section, we
are provided with a set of $M$ images with which we must texture our
target plane. Each image has a camera matrix $P_i$ for $i=1..M$, which
translates a 3D point in the world coordinate system to a 2D point or
pixel in image $i$'s coordinates. If the 3D world point is not
contained in the image, the 2D point will simply be outside of the
image boundaries. A camera matrix $P_i$ is composed of the camera's
intrinsic parameters, such as focal length and image center, as well
as extrinsic parameters which specify the rotation and translation of
the camera's position in 3D world coordinates at the time that image
$i$ was taken. These extrinsic parameters are determined by the
backpack hardware and localization algorithms mentioned in Section
2. A point $X$ on the plane in 3D space can be related to its
corresponding pixel $x$ in image $i$ through the following equation:

@@

Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
we can texture the plane by discretizing it into small square tiles,
generally about 5 pixels across, and picking an image to texture each
tile with. Since we are provided with far more images than necessary
to cover our plane, this process is a simple way of assigning
texturess to locations on our plane. We choose to work with
rectangular units to ensure that borders between any two distinct
images in our final texture will be either horizontal or
vertical. Since most environmental features inside buildings are
horizontal or vertical, any seams in our texture will intersect them
miniimally and be less noticeable.

In order to select an image for texturing tile $t$, we must first
gather a list of candidate images that contain all four of its
corners, which we can quickly check by projecting $t$ into each image
and using the projection method above. Furthermore, each candidate
image must have been taken at a time when its camera had a clear
line-of-sight to $t$, which can be calculated using
standard ray-polygon intersection tests between the camera location,
the center of $t$, and other planes, all in world coordinates.

@@

Similar to a caching mechanism, we select the best image for a tile
$t$ by searching through two subsets of images for a good candidate,
before searching through the entire set. The first subset of images is
the set of images selected by adjacent tiles that have already been
processed. We must first check which images can map to $t$, and then
of those, we make a choice according to the same scoring function in
figure \ref{fig:scoringFunction}. Rather than blindly reusing this
image, we ensure it meets a predefined score threshold to be
considered a good image. If no good image is found, we then check our
second subset of images, which consists of images that were taken near
the images in the first subset, both spatially and temporally. These
images are not the same as the ones used for neighboring tiles, but
they were taken at a similar location and time, suggesting that their
localization and projection will be very similar. Again, if no good
image is found according to a threshold, we then must search the
entire set of candidate images.

@@

The results of this caching approach are shown in figure
\ref{fig:caching}.  Discontinuities have been reduced overall, but the
amount of remaining seams suggests that image selection alone cannot
produce seamless textures. The image projections themselves will have
to be adjusted in order to reliably generate clean textures.

@@

Rather than perform camera or image adjustments across many thousands
of images, we choose to work with the more limited set of images
corresponding to those chosen by the direct mapping approach, without
caching. This set of images constitutes a good candidate set for
generating a final seamless texture, since it meets three important
criteria. First, this set of images contains at least one image that
covers each tile on our plane. Thus, unless extreme changes in
localization occur, we can ensure that there will be no holes in our
final texture. Second, since images are all selected according to the
same scoring function in figure \ref{fig:scoringFunction}, we know
that our images are all taken at as much of a head-on angle as
possible and should project onto the plane in similar ways. Third, as
a side result of the scoring function, selected images are only good
candidates for the tiles near their center of projection. Thus, there
should be plenty of overlap between selected images, allowing for some
degree of shifting without resulting in holes, as well as area for
blending between them. With this set of images, we now examine two
approaches towards refining and combining their projections, before
demonstrating our own.

@@

It should be evident that our localization adjustments resulted in a
major improvement. A great strength of our tiling approach for image
selection is that it extracts the areas with good camera angles from
each image, and thus can handle images of all types. As demonstrated
on the floor plane in figure \ref{fig:floor_suboptimal}, we can run
this process even with extremely poor image projections, where the
amount of quality texture for each image is limited. Additionally,
since our plane is discretized, we can handle arbitrarily
complex plane geometry as well.

@@

\subsection{Texture Mapping with Seam Minimization}
The cached tile approach selected images based on individual and
neighboring quality of image selections, in an effort to reduce seams
between tiles. In a more optimal case, as the backpack system provides
for walls, images are all taken from close distances and have
near-rectangular projections and thus more similar scores according to the
scoring function in figure \ref{fig:scoringFunction}. Thus, we can
reason that rather than selecting the set of best images, since all
images are good, we should instead select the best set of images, such
that the selection together results in the cleanest final texture. We
will accomplish this by using entire images where possible, defining a
cost function between images, and seeking to minimize total cost.

@@

Fortunately, by virtue of our indoor environments, the vast majority
of surface geometry is either horizontal or vertical, with high
amounts of right angles. This means that after masking out occluded
areas, our image projections will remain largely rectangular. We can
thus make efficiency improvements by recursively splitting each image
into rectangular pieces, and performing occlusion checks where
needed. To actually occlude out rectangles, we simply remove their
texture, as we will ensure that untextured areas are never chosen for
texture mapping.

\subsubsection{Inter-Image Cost Functions}
In order to objectively decide which set of images results in the
cleanest texture, we need a cost function such that we can evaluate
the visibiilty of seams between images in our set. A straightforward
cost function that accomplishes this is the sum of squared pixel
differences in overlapping regions between all pairs of
images. Minimizing this cost function encourages image boundaries to
occur in featureless areas, such as bare walls, as well as in areas
where images match extremely well.

Another possible cost function is overall edge energy, i.e. the sum of
the smoothed gradient over seams. Minimizing this encourages image
boundaries to be placed in featureless areas even more than the first
cost function. For the results shown throughout this paper, the first
cost function was used, as the regular occurrence of featureless areas
was not guaranteed in our datasets.

@@

Now that we have a cost function defined, we mechanically select the
set of images for which the overall cost function is minimized. Since
we aim to cover the entirety of our plane, our problem is to minimally
cover a polygon(our plane), using other polygons of arbitrary geometry
(our image projections), with the added constraint of minimizing our
cost function between chosen images. This is a complex problem, though
we can take a number of steps to simplify it. Returning again to the
optimality of our situation when texturing walls, we can make a quick
simplification for the sake of efficiency. Given that our wall-texture
candidate images were all taken from a head-on angle, and assuming
only minor rotations were made during localization refinement, we can
reason that their projections onto the plane are approximately
rectangular. By discarding the minor excess texture and cropping them
all to be rectangular, our problem becomes the conceptually simpler
problem of filling a polygon with rectangles, such that the sum of all
edge costs between each pair of rectangles is minimal. We thus also
retain axis-aligned image boundaries, along with their advantages
explained in section 3.

By taking two more shortcut we can simplify our problem down even
further. Since this approach is already assuming optimal images, we
know it will be applied on walls, which are the focus of the backpack
modeling system. For wall planes, which are nearly always rectangular,
care is taken so that images contain the entirety of their
floor-to-ceiling range, and thus we do not have wall images such that
one should be projected vertically above the other. In essence, we
need only to ensure horizontal coverage of our planes, as our images
provide full vertical coverage themselves. We can thus construct a
Directed Acyclic Graph (DAG) from the images, with edge costs defined
by our cost function above, and solve a simple shortest path problem
to find an optimal subset of images with regard to the cost functions.

@@

In rare cases where the vertical dimension of the plane is not
entirely covered by a chosen image, we are left with a hole where no
image was chosen to texture. Rather than reverting to a 2D-coverage
problem, we can elect to simply fill the hole by selecting images to
fill it in a greedy fashion with respect to edge costs of the same
cost function.

@@

Although our preprocessing steps and image selections in either method
attempt to minimize all mismatches between images, there are cases,
where due to different lighting conditions or inaccuracies in geometry
or projection, where we simply have unavoidable discontinuities. These
can however be treated and smoothed over by applying alpha blending
over image seams.  Whether the units we are blending are
rectangularly-cropped images or rectangular tiles, we can apply the
same blending procedure, as long as we have a guaranteed overlap
between units to blend over.

@@

\section{Conclusion and Results}
In this paper, we have shown how to accurately texture map models,
even when provided with imprecise camera localization data. We were
able to refine image locations based on feature mapping, and robustly
handle outliers.  We generalized one approach to texture mapping to
any manner of planes and images, and successfully textured both simple
rectangular walls as well as complex floor and ceiling geometry. We
also presented an optimized texturing method that takes advantage of
our localization refinement process and produces cleaner textures on
planes where multiple head-on images are available. Each of these
approaches is highly modular, and easily tunable for different
environments and acquisition hardware.

Ceilings and floors textured
with the robust approach, and walls textured with the optimized
approach, are displayed in figure \ref{fig:results}. A more detailed
walkthrough demonstrating fully textured 3D models using the
approaches in this paper is available in the accompanying video to
this paper.
